TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-09-20 paper/draft.tex
completed bullet
rewrite abstract bits

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution, near-infrared \apogee\ spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H], and detailed abundances) to precisions higher than the \apogee\ survey model-grid pipelines.
The lack of reliable stellar models in the near infrared for temperatures $<3500$\,K motivates the extension of this approach to M-dwarf stars, but with new training labels from optical surveys that overlap \sdssiv\ \apogee.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex respectively.
These are parameter-space precision statements, but the model is accurate in the data space: It accurately predicts \apogee\ spectral pixels at the percent level.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging from M0 to M9, and demonstrate that \thecannon\ acting on \apogee\ spectra can infer spectral types with predictive accuracy of 0.9 types.
The trained models are employed to determine spectral subtypes, temperatures and metallicities for 10,311 M dwarfs in \apogee, selected in part on \gaia\ photometry and astrometry.
We validate our results using photometric temperature indicators and the \gaia\ color--magnitude diagram.
As a side result, we determine empirical relationships between spectral type and temperature and metallicity.
These results have implications for the NASA \tess\ Mission and other projects that involve large samples of M dwarfs.

@@

JB ALSO POINT TO FIGURES THAT CONNECT TYPES TO PARAMETERS. AND DISCUSS.

JB DO WE WANT A FIGURE THAT SHOWS ABUNDANCES AS A FUNCTION OF MILKY WAY POSITION OR ACTIONS?

>>>>

\begin{abstract}
\thecannon\ is a flexible, data-driven spectral modeling and parameter inference framework, demonstrated on high-resolution, near-infrared \apogee\ spectra of giant stars to estimate stellar labels (\teff, logg, [Fe/H], and detailed abundances) to precisions higher than the \apogee\ survey model-grid pipelines.
The lack of reliable atmospheric models in the near infrared for temperatures cooler than $\sim3500$\,K motivates the extension of this approach to M-dwarf stars, but with calibrated training labels from optical surveys that overlap \sdssiv\ \apogee.
Using a training sample of 87 M dwarfs in the temperature range $2859 < T_{\rm eff} < 4131$\,K, and metallicty range $-0.48 < \mathrm{[Fe/H]}< 0.49$\,dex, we train a two-parameter model with predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex respectively.
These are parameter-space precision statements, but the model is accurate in the data space: it accurately predicts \apogee\ spectral pixels at the percent level.
We also train a one-dimensional spectral type model using 51 M dwarfs with known spectral types ranging from M0 to M9, with predictive accuracy of 0.9 types. From the \textsl{APOGEE DR14} -- \textsl{Gaia DR2} overlap we select a test sample and employ our trained models to determine temperatures, metallicities and spectral subtypes for nearly all of the M dwarfs in \apogee\ (10,311 sources), and validate our \cannon\ derived \teff\ against color-temperature relations and [Fe/H] against isochrone models in \gaia\ color-magnitude space. Finally we use 
These results have implications for the NASA \tess\ Mission and other projects that involve large samples of M dwarfs.

% The trained models are employed to determine spectral subtypes, temperatures and metallicities for 10,311 M dwarfs in \apogee, selected in part on \gaia\ photometry and astrometry.
% We validate our results using photometric temperature indicators and the \gaia\ color--magnitude diagram.
% As a side result, we determine empirical relationships between spectral type and temperature and metallicity.

@@

% JB DO WE WANT A FIGURE THAT SHOWS ABUNDANCES AS A FUNCTION OF MILKY WAY POSITION OR ACTIONS?
